---
title: "Example rOpenSci Packages"
author: "Martin Fenner"
date: "5. Oktober 2015"
output:
  html_document:
    keep_md: true
---

[rOpenSci](https://ropensci.org) provides R packages for scholarly APIs, including altmetrics services.

### alm
A package for the [lagotto](https://github.com/lagotto/lagotto) open source application. The [alm tutorial](https://ropensci.org/tutorials/alm_tutorial.html) is a good starting point.

Install and load the package with 
```{r}
#install.packages("alm")
library(alm)
```

My default the package talks to the lagotto instance run by the publisher PLOS at `http://alm.plos.org`. You can specify the `url` parameter to talk to a different instance of the Lagotto API, e.g. `http://cls.labs.datacite.org`.

Get metrics for a particular DOI
```{r}
out <- alm_ids(doi="10.1371/journal.pone.0029797")
out$data
```

get metrics for a set of DOIs
```{r}
dois <- c('10.1371/journal.pone.0001543','10.1371/journal.pone.0040117',
    '10.1371/journal.pone.0029797','10.1371/journal.pone.0039395')
out <- alm_ids(doi=dois)
lapply(out$data, head)
```